\section{Response Tracker}
\subsection{Overview}

The Response Tracker is a write-path block responsible for aggregating
per-packet MC Core completions into a single AXI B channel response per
transaction. It tracks every outstanding write transaction from tag
allocation to final BRESP emission.

Its responsibilities are:
\begin{itemize}
  \item Store per-transaction metadata at allocation time: expected
        packet count, completed packet count, error status.
  \item Increment the completed packet counter on each \texttt{WR\_ACK}
        completion received from MC Core.
  \item Assert \texttt{txn\_done} when all expected packets for a
        transaction have completed.
  \item Enforce AXI4 B channel ordering: BRESP is emitted in-order per
        AXID using a per-AXID head pointer.
  \item Propagate error status: if any packet in a burst returns an
        error completion, the sticky error bit is set and the
        transaction generates \texttt{BRESP = DECERR}.
  \item Free the occupied Tag Allocator slot on BRESP emission via the
        \texttt{WR\_FREE} port.
\end{itemize}

\textbf{Note on abstraction levels.} The full CIF pipeline diagram
shows MC Core as the sole input to the Response Tracker. At the
internal logic level, the Tag Allocator also provides an allocation
signal — this is a resource management path (writing \texttt{exp\_pkts}
at tag assignment time) that is separate from the data completion path.
Both are correct; they operate at different abstraction levels.

% -----------------------------------------------------------------------
\subsection{Design Parameters}

\begin{table}[h!]
\centering
\begin{tabular}{lll}
\toprule
\textbf{Parameter} & \textbf{Value} & \textbf{Description} \\
\midrule
\texttt{AXID\_W}          & 6   & AXI ID width; 64 unique IDs \\
\texttt{SEQNUM\_W}        & 2   & Seqnum width \\
\texttt{TXN\_TAG\_W}      & 8   & Tag width = AXID\_W + SEQNUM\_W \\
\texttt{RT\_DEPTH}        & 256 & Register file entries = $2^8$ \\
\texttt{EXP\_PKTS\_W}     & 8   & Expected packets width (covers AWLEN+1, max 256) \\
\texttt{CMP\_PKTS\_W}     & 8   & Completed packets counter width \\
\texttt{N\_AXIDS}         & 64  & Head pointer array depth \\
\texttt{HEAD\_W}          & 8   & Head pointer width = TXN\_TAG\_W \\
\bottomrule
\end{tabular}
\caption{Response Tracker parameters}
\end{table}

% -----------------------------------------------------------------------
\subsection{Internal Blocks}

\subsection{Register File}

\begin{itemize}
  \item \textbf{Structure:} 256-entry register file. Each entry stores:
  \begin{itemize}
    \item \texttt{exp\_pkts[7:0]} — expected packet count (= \texttt{AWLEN+1})
    \item \texttt{cmp\_pkts[7:0]} — completed packet count
    \item \texttt{status[3:0]} — MC Core status code from most recent
          error completion
    \item \texttt{err\_sticky[0]} — sticky error bit; set on first
          error completion, never cleared until entry is freed
  \end{itemize}
  \item \textbf{Addressing:} \texttt{addr[7:0] = TXN\_TAG[7:0] =
        \{AXID[5:0], seqnum[1:0]\}}. Tag directly indexes the register
        file — no translation required.
  \item \textbf{Write ports:}
  \begin{itemize}
    \item Alloc Write Logic: writes \texttt{exp\_pkts}, initialises
          \texttt{cmp\_pkts = 0}, \texttt{err\_sticky = 0} on
          allocation.
    \item Completion Update Logic: increments \texttt{cmp\_pkts},
          optionally sets \texttt{err\_sticky} and \texttt{status} on
          each \texttt{WR\_ACK} completion.
    \item Free Logic: clears the entry on BRESP emission.
  \end{itemize}
  \item \textbf{Read ports:} Continuously exposes \texttt{cmp\_pkts}
        and \texttt{exp\_pkts} to Compare Logic; exposes
        \texttt{err\_sticky} and \texttt{status} to Emit Logic.
\end{itemize}

\subsection{Alloc Write Logic}

\begin{itemize}
  \item Activated on \texttt{TAG\_VALID} from the Tag Allocator.
  \item Receives \texttt{TXN\_TAG[7:0]} and \texttt{AWLEN[7:0]}.
  \item Computes \texttt{exp\_pkts = AWLEN + 1}.
  \item Issues \texttt{write\_en} to the Register File at address
        \texttt{TXN\_TAG[7:0]}, writing \texttt{exp\_pkts} and
        initialising \texttt{cmp\_pkts = 0}, \texttt{err\_sticky = 0}.
\end{itemize}

\subsection{Completion Update Logic}

\begin{itemize}
  \item Taps the MC Core completion bus: \texttt{RESP\_VALID},
        \texttt{resp\_type[1:0]}, \texttt{tag[7:0]},
        \texttt{status[3:0]}.
  \item \textbf{Filter:} acts only when \texttt{resp\_type == 2'b01}
        (WR\_ACK). RD\_DATA completions (\texttt{resp\_type == 2'b00})
        are destined for the ROB and RTT and are ignored here.
  \item On a normal WR\_ACK (\texttt{resp\_type == 2'b01}, no error):
        asserts \texttt{inc\_en} to the Register File at address
        \texttt{tag[7:0]}, incrementing \texttt{cmp\_pkts} by 1.
  \item On an error completion (\texttt{resp\_type == 2'b10}):
        increments \texttt{cmp\_pkts} and additionally sets
        \texttt{err\_sticky = 1} and latches \texttt{status[3:0]}.
        The sticky bit is never cleared by subsequent completions —
        one error poisons the entire transaction.
\end{itemize}

\subsection{Compare Logic}

\begin{itemize}
  \item Purely combinational.
  \item Reads \texttt{cmp\_pkts[7:0]} and \texttt{exp\_pkts[7:0]} from
        the Register File for the entry currently being updated.
  \item Asserts \texttt{txn\_done} as a one-cycle pulse when
        \texttt{cmp\_pkts == exp\_pkts}.
  \item \texttt{txn\_done} is forwarded to Emit Logic to trigger BRESP
        evaluation.
\end{itemize}

\subsection{Head Pointer Array}

\begin{itemize}
  \item \textbf{Structure:} 64 × 8-bit registers, one per AXID.
        Each register holds the \texttt{TXN\_TAG} of the next
        transaction expected to emit BRESP for that AXID.
  \item \textbf{Init:} Written on first allocation for an AXID
        (occupancy 0→1). On subsequent allocations for the same AXID,
        the head pointer is not updated — it is owned by Emit Logic
        from that point forward.
  \item \textbf{Advance:} Emit Logic increments the head pointer
        (mod 4 on the seqnum field) after each BRESP is emitted for
        that AXID.
  \item \textbf{Purpose:} Enforces AXI4 B channel ordering — BRESP
        is only emitted for the transaction at the head, not for
        any later-completing transaction of the same AXID.
\end{itemize}

\subsection{Emit Logic}

\begin{itemize}
  \item Checks whether the transaction at \texttt{head\_tag[AXID]} is
        done: reads \texttt{txn\_done} from Compare Logic and confirms
        the completing tag matches the current head pointer for that
        AXID.
  \item If head slot is done:
  \begin{itemize}
    \item Reads \texttt{err\_sticky} and \texttt{status} from the
          Register File.
    \item Asserts \texttt{BRESP\_VALID} to the B Response Generator.
    \item Drives \texttt{BID[5:0] = head\_tag[7:2]} (AXID recovered
          from tag).
    \item Drives \texttt{BRESP[1:0] = OKAY} if \texttt{err\_sticky
          == 0}; \texttt{DECERR} if \texttt{err\_sticky == 1}.
    \item Asserts \texttt{advance\_head} to the Head Pointer Array
          for the relevant AXID.
    \item Asserts \texttt{emit\_done} with \texttt{tag[7:0]} to Free
          Logic.
  \end{itemize}
  \item If head slot is not yet done: Emit Logic holds. Later-completing
        transactions for the same AXID are buffered in the Register File
        until the head slot completes.
\end{itemize}

\subsection{Free Logic}

\begin{itemize}
  \item Activated by \texttt{emit\_done} from Emit Logic.
  \item Asserts \texttt{clear\_en} to the Register File at address
        \texttt{tag[7:0]}, clearing the entry in the same cycle as
        BRESP emission.
  \item Simultaneously asserts \texttt{WR\_FREE\_VALID} and drives
        \texttt{WR\_FREE\_TAG[7:0]} to the Tag Allocator's WR\_FREE
        port, decrementing the occupancy counter for the freed AXID.
\end{itemize}

% -----------------------------------------------------------------------
\subsection{Interface Signals}

\subsection{Inputs}

\begin{longtable}{p{3cm}p{1cm}p{3cm}p{7cm}}
\toprule
\textbf{Signal} & \textbf{Width} & \textbf{Source} & \textbf{Description} \\
\midrule
\texttt{TXN\_TAG[7:0]}     & 8 & Tag Allocator & Allocated tag at write transaction time \\
\texttt{AWLEN[7:0]}        & 8 & Tag Allocator & Burst length; \texttt{exp\_pkts = AWLEN+1} \\
\texttt{TAG\_VALID}        & 1 & Tag Allocator & Allocation strobe \\
\texttt{RESP\_VALID}       & 1 & MC Core       & Completion valid \\
\texttt{resp\_type[1:0]}   & 2 & MC Core       & 00=RD\_DATA (ignored), 01=WR\_ACK, 10=ERR \\
\texttt{tag[7:0]}          & 8 & MC Core       & Completion tag; used as register file address \\
\texttt{status[3:0]}       & 4 & MC Core       & Error status; latched on error completion \\
\bottomrule
\caption{Response Tracker input signals}
\end{longtable}

\subsection{Outputs}

\begin{longtable}{p{3cm}p{1cm}p{4cm}p{5.5cm}}
\toprule
\textbf{Signal} & \textbf{Width} & \textbf{Destination} & \textbf{Description} \\
\midrule
\texttt{BRESP\_VALID}      & 1 & B Response Generator          & BRESP ready to emit \\
\texttt{BID[5:0]}          & 6 & B Response Generator          & AXI ID recovered from \texttt{head\_tag[7:2]} \\
\texttt{BRESP[1:0]}        & 2 & B Response Generator          & OKAY or DECERR \\
\texttt{WR\_FREE\_VALID}   & 1 & Tag Allocator (WR\_FREE port) & Free strobe; coincident with BRESP emit \\
\texttt{WR\_FREE\_TAG[7:0]}& 8 & Tag Allocator (WR\_FREE port) & Tag being freed \\
\bottomrule
\caption{Response Tracker output signals}
\end{longtable}

\subsection{Internal Signals}

\begin{longtable}{p{2.8cm}p{1cm}p{2.7cm}p{2.7cm}p{5cm}}
\toprule
\textbf{Signal} & \textbf{Width} & \textbf{From} & \textbf{To} & \textbf{Description} \\
\midrule
\texttt{write\_en}          & 1   & Alloc Write Logic       & Register File       & Write strobe on allocation \\
\texttt{wr\_addr[7:0]}      & 8   & Alloc Write Logic       & Register File       & = \texttt{TXN\_TAG[7:0]} \\
\texttt{exp\_pkts[7:0]}     & 8   & Alloc Write Logic       & Register File       & = \texttt{AWLEN+1} \\
\texttt{inc\_en}            & 1   & Completion Update Logic & Register File       & Increment \texttt{cmp\_pkts} strobe \\
\texttt{err\_set}           & 1   & Completion Update Logic & Register File       & Set \texttt{err\_sticky} on error \\
\texttt{inc\_addr[7:0]}     & 8   & Completion Update Logic & Register File       & = \texttt{tag[7:0]} from MC Core \\
\texttt{cmp\_pkts[7:0]}     & 8   & Register File           & Compare Logic       & Current completed count \\
\texttt{exp\_pkts[7:0]}     & 8   & Register File           & Compare Logic       & Expected count (read port) \\
\texttt{err\_sticky}        & 1   & Register File           & Emit Logic          & Sticky error flag \\
\texttt{txn\_done}          & 1   & Compare Logic           & Emit Logic          & Pulse: all packets completed \\
\texttt{head\_tag[7:0]}     & 8   & Head Pointer Array      & Emit Logic          & Next tag to emit per AXID \\
\texttt{advance\_head}      & 1   & Emit Logic              & Head Pointer Array  & Increment head pointer for AXID \\
\texttt{emit\_done}         & 1   & Emit Logic              & Free Logic          & BRESP emitted; trigger free \\
\texttt{clear\_en}          & 1   & Free Logic              & Register File       & Clear entry strobe \\
\texttt{clear\_addr[7:0]}   & 8   & Free Logic              & Register File       & = tag of emitted transaction \\
\bottomrule
\caption{Response Tracker internal signals}
\end{longtable}

% -----------------------------------------------------------------------
\subsection{Functional Behaviour}

\subsection{Allocation Flow}

\begin{enumerate}
  \item Tag Allocator asserts \texttt{TAG\_VALID} with
        \texttt{TXN\_TAG[7:0]} and \texttt{AWLEN[7:0]}.
  \item Alloc Write Logic computes \texttt{exp\_pkts = AWLEN + 1}.
  \item Register File is written at address \texttt{TXN\_TAG[7:0]}:
        \texttt{exp\_pkts} stored, \texttt{cmp\_pkts = 0},
        \texttt{err\_sticky = 0}.
  \item If this is the first in-flight transaction for this AXID
        (occupancy 0→1), Head Pointer Array is initialised with
        \texttt{TXN\_TAG[7:0]}.
\end{enumerate}

\subsection{Completion Flow}

\begin{enumerate}
  \item MC Core asserts \texttt{RESP\_VALID} with
        \texttt{resp\_type == 2'b01} (WR\_ACK) and \texttt{tag[7:0]}.
  \item Completion Update Logic asserts \texttt{inc\_en} at address
        \texttt{tag[7:0]}: \texttt{cmp\_pkts} incremented by 1.
  \item If \texttt{resp\_type == 2'b10} (ERR): additionally sets
        \texttt{err\_sticky = 1} and latches \texttt{status[3:0]}.
  \item Compare Logic evaluates \texttt{cmp\_pkts == exp\_pkts}
        combinationally; asserts \texttt{txn\_done} if true.
\end{enumerate}

\subsection{Emit Flow}

\begin{enumerate}
  \item \texttt{txn\_done} pulses for a completing transaction.
  \item Emit Logic checks whether the completing tag matches
        \texttt{head\_tag} for that AXID.
  \item If match (in-order): reads \texttt{err\_sticky} from Register
        File, drives \texttt{BRESP\_VALID}, \texttt{BID}, and
        \texttt{BRESP} (OKAY or DECERR) to B Response Generator.
        Asserts \texttt{advance\_head} and \texttt{emit\_done}.
  \item If no match (out-of-order): transaction is complete in the
        Register File but the head has not yet completed. Emit Logic
        holds. BRESP is deferred until the head slot completes.
\end{enumerate}

\subsection{Free Flow}

\begin{enumerate}
  \item \texttt{emit\_done} triggers Free Logic in the same cycle as
        BRESP emission.
  \item Free Logic asserts \texttt{clear\_en} to Register File:
        entry is invalidated.
  \item \texttt{WR\_FREE\_VALID} and \texttt{WR\_FREE\_TAG} are
        asserted to Tag Allocator, decrementing occupancy for the
        freed AXID.
\end{enumerate}

% -----------------------------------------------------------------------
\subsection{Timing Notes}

\begin{itemize}
  \item Alloc Write Logic is registered: Register File write takes
        effect the cycle after \texttt{TAG\_VALID}.
  \item Completion Update Logic is registered: \texttt{cmp\_pkts}
        increment takes effect the cycle after \texttt{RESP\_VALID}.
  \item Compare Logic is combinational: \texttt{txn\_done} appears in
        the same cycle the increment is visible — one cycle after the
        final \texttt{RESP\_VALID}.
  \item Emit Logic is registered: \texttt{BRESP\_VALID} is asserted
        the cycle after \texttt{txn\_done} is evaluated.
  \item Free Logic fires in the same cycle as BRESP emission:
        \texttt{clear\_en} and \texttt{WR\_FREE\_VALID} are asserted
        without additional latency.
  \item Simultaneous allocation and completion on the same tag are not
        possible by construction — a tag is only re-allocated after its
        slot is freed, which occurs after BRESP emission.
\end{itemize}

% -----------------------------------------------------------------------
\subsection{Edge Cases}

\begin{itemize}
  \item \textbf{Single-packet transaction (AWLEN = 0):}
        \texttt{exp\_pkts = 1}. \texttt{txn\_done} fires on the first
        and only WR\_ACK. Handled identically to multi-packet
        transactions.
  \item \textbf{Partial error in a burst:} If one packet in a burst
        returns ERR and others return WR\_ACK, \texttt{err\_sticky} is
        set on the first error. All subsequent completions still
        increment \texttt{cmp\_pkts} normally. On BRESP emission,
        \texttt{BRESP = DECERR} regardless of how many packets
        succeeded.
  \item \textbf{Out-of-order completion within an AXID:} AXI4 requires
        B channel responses to be in-order per AXID. If a later
        transaction of an AXID completes before the head transaction,
        Emit Logic holds the BRESP until the head slot is done. The
        later transaction's completion is safely held in the Register
        File.
  \item \textbf{RD\_DATA completions:} Ignored entirely by Completion
        Update Logic. These are processed by the ROB and RTT on the
        read path.
  \item \textbf{Over-completion protection:} \texttt{cmp\_pkts} must
        never exceed \texttt{exp\_pkts}. Guaranteed by construction
        since MC Core returns exactly one WR\_ACK per issued write
        packet. A design-time assertion should be included in the
        verification environment.
  \item \textbf{Simultaneous alloc and completion on same AXID
        (different seqnums):} These target different register file
        entries (different \texttt{tag[7:0]} addresses) — no conflict.
\end{itemize}