\documentclass[11pt,a4paper]{article}

\usepackage[a4paper, margin=2.5cm]{geometry}
\usepackage{booktabs}
\usepackage{array}
\usepackage{longtable}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{parskip}
\usepackage{fancyhdr}
\usepackage{listings}

\definecolor{codegray}{rgb}{0.95,0.95,0.95}
\lstset{
  backgroundcolor=\color{codegray},
  basicstyle=\ttfamily\small,
  frame=single,
  framesep=4pt,
  breaklines=true
}

\hypersetup{colorlinks=true, linkcolor=blue, urlcolor=blue}

\pagestyle{fancy}
\fancyhf{}
\rhead{CIF Microarchitecture}
\lhead{Request Validator}
\rfoot{\thepage}

\title{\textbf{CIF Microarchitecture Specification}\\[0.4em]
\large Block: Request Validator}
\author{CIF Design Team}
\date{\today}

\begin{document}

\maketitle
\thispagestyle{fancy}
\tableofcontents
\newpage

% -----------------------------------------------------------------------
\section{Overview}

The Request Validator is a purely combinational block instantiated once
each on the read and write paths. It pops entries from the AR or AW
Metadata FIFO, performs a single runtime check (address range
validation), and steers the transaction to one of two downstream paths:

\begin{itemize}
  \item \textbf{Valid path:} Transaction forwarded to the Transaction
        Tag Allocator with \texttt{REQ\_VALID} asserted.
  \item \textbf{Error path:} Transaction forwarded to the Error Handler
        with \texttt{ERR\_VALID} asserted and \texttt{AXID} preserved.
\end{itemize}

The Request Validator contains no storage, no counters, and no state.
All protocol-level checks (burst type, WRAP legality, alignment, burst
length) are enforced as design-time assertions in the verification
environment only. Compliant AXI4 masters are assumed at runtime.

% -----------------------------------------------------------------------
\section{Design Parameters}

\begin{table}[h!]
\centering
\begin{tabular}{lll}
\toprule
\textbf{Parameter} & \textbf{Value} & \textbf{Description} \\
\midrule
\texttt{ADDR\_W}  & 32 & Address width in bits \\
\texttt{AXID\_W}  & 6  & AXI ID width \\
\texttt{LEN\_W}   & 8  & Burst length field width (ARLEN/AWLEN) \\
\bottomrule
\end{tabular}
\caption{Request Validator parameters}
\end{table}

% -----------------------------------------------------------------------
\section{Internal Blocks}

\subsection{Address Range Check}

\begin{itemize}
  \item \textbf{Type:} Purely combinational.
  \item \textbf{Input:} \texttt{ADDR[31:0]} from the AR/AW Metadata
        FIFO.
  \item \textbf{Function:} Checks whether \texttt{ADDR} falls within
        the valid memory address range defined at synthesis time.
  \item \textbf{Output:} \texttt{addr\_ok} (1-bit) — asserted when the
        address is valid, deasserted on an address error.
  \item \textbf{Scope:} Address range check only. No burst type, no
        alignment, no length checks at runtime.
\end{itemize}

\subsection{Output Mux}

\begin{itemize}
  \item \textbf{Type:} Purely combinational.
  \item \textbf{Function:} Routes the transaction to the correct
        downstream block based on \texttt{addr\_ok}.
  \item \textbf{Valid path} (\texttt{addr\_ok == 1}): asserts
        \texttt{REQ\_VALID} to the Tag Allocator; forwards
        \texttt{AXID[5:0]} and \texttt{ARLEN/AWLEN[7:0]}.
  \item \textbf{Error path} (\texttt{addr\_ok == 0}): asserts
        \texttt{ERR\_VALID} to the Error Handler; forwards
        \texttt{AXID[5:0]}.
  \item Only one output path is active per transaction. Both paths are
        never asserted simultaneously.
\end{itemize}

% -----------------------------------------------------------------------
\section{Interface Signals}

\subsection{Inputs}

\begin{longtable}{llll}
\toprule
\textbf{Signal} & \textbf{Width} & \textbf{Source} & \textbf{Description} \\
\midrule
\texttt{ADDR[31:0]}        & 32 & AR/AW Metadata FIFO & Transaction address \\
\texttt{AXID[5:0]}         & 6  & AR/AW Metadata FIFO & AXI ID \\
\texttt{ARLEN/AWLEN[7:0]}  & 8  & AR/AW Metadata FIFO & Burst length \\
\texttt{STALL}             & 1  & Tag Allocator       & Per-AXID stall; holds output until deasserted \\
\bottomrule
\caption{Request Validator input signals}
\end{longtable}

\subsection{Outputs}

\begin{longtable}{llll}
\toprule
\textbf{Signal} & \textbf{Width} & \textbf{Destination} & \textbf{Description} \\
\midrule
\texttt{REQ\_VALID}       & 1 & Tag Allocator & Valid transaction request \\
\texttt{AXID[5:0]}        & 6 & Tag Allocator & Forwarded AXI ID \\
\texttt{ARLEN/AWLEN[7:0]} & 8 & Tag Allocator & Forwarded burst length \\
\texttt{ERR\_VALID}       & 1 & Error Handler & Address error notification \\
\texttt{AXID[5:0]}        & 6 & Error Handler & AXI ID preserved for NOP tag allocation \\
\bottomrule
\caption{Request Validator output signals}
\end{longtable}

% -----------------------------------------------------------------------
\section{Functional Behaviour}

\begin{enumerate}
  \item Request Validator pops the head entry from the AR/AW Metadata
        FIFO: \texttt{ADDR[31:0]}, \texttt{AXID[5:0]},
        \texttt{ARLEN/AWLEN[7:0]}.
  \item Address Range Check evaluates \texttt{addr\_ok} combinationally.
  \item If \texttt{addr\_ok == 1} and \texttt{STALL} is deasserted:
        Output Mux asserts \texttt{REQ\_VALID} to the Tag Allocator
        with \texttt{AXID} and \texttt{ARLEN/AWLEN}.
  \item If \texttt{addr\_ok == 0}: Output Mux asserts \texttt{ERR\_VALID}
        to the Error Handler with \texttt{AXID}. The Error Handler sets
        \texttt{pkt\_type = Err\_write / Err\_read} and injects the
        transaction into the normal pipeline via the Tag Allocator.
  \item If \texttt{STALL} is asserted (Tag Allocator full for this
        AXID): the Request Validator holds its outputs stable and does
        not pop the next FIFO entry until \texttt{STALL} deasserts.
\end{enumerate}

% -----------------------------------------------------------------------
\section{Timing Notes}

\begin{itemize}
  \item Fully combinational. Output (\texttt{REQ\_VALID} or
        \texttt{ERR\_VALID}) is valid in the same cycle as the FIFO
        head entry is presented.
  \item \texttt{STALL} from the Tag Allocator is also combinational —
        the combined path (FIFO output → Address Range Check → Output
        Mux → Tag Allocator Stall Logic → STALL back) is a
        combinational loop that must be timed carefully. A register
        stage may be inserted between the Request Validator output and
        the Tag Allocator input if timing closure requires it, at the
        cost of one cycle of allocation latency.
\end{itemize}

% -----------------------------------------------------------------------
\section{Edge Cases}

\begin{itemize}
  \item \textbf{Error path stall:} An address error transaction still
        consumes a Tag Allocator slot. If the AXID is at maximum
        outstanding (\texttt{STALL} asserted), the error path is also
        held. \texttt{ERR\_VALID} is not asserted until \texttt{STALL}
        deasserts.
  \item \textbf{Protocol violations:} Illegal burst type, unaligned
        WRAP, and maximum burst length violations are not checked at
        runtime. They are enforced as design-time assertions in the
        verification environment. A non-compliant master will cause
        undefined behaviour.
  \item \textbf{Simultaneous valid and error:} Not possible —
        \texttt{addr\_ok} is a single-bit result; exactly one output
        path is active per transaction.
\end{itemize}

\end{document}
