\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{array}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{titlesec}
\usepackage{parskip}
\usepackage{microtype}
\usepackage{hyperref}
\usepackage{enumitem}

\hypersetup{colorlinks=true, linkcolor=black, urlcolor=blue}

\titleformat{\section}{\large\bfseries}{}{0em}{}[\titlerule]
\titleformat{\subsection}{\normalsize\bfseries}{}{0em}{}

\title{\textbf{ROB Internal Logic}}
\date{}
\author{}

\begin{document}
\maketitle
\thispagestyle{empty}

\section{Overview}

The Reorder Buffer (ROB) absorbs out-of-order read completions from MC Core and emits them in-order per AXID to Merge Logic. Depth is tied to \texttt{MC\_RD\_BUF = 64}. Since \texttt{RESP\_READY} must always be held high, the ROB can never stall incoming completions — backpressure is instead applied by halting new read requests when occupancy reaches the watermark threshold.

\subsection*{Input Signals}
\begin{longtable}{llll}
\toprule
\textbf{Signal} & \textbf{Width} & \textbf{Source} & \textbf{Description} \\
\midrule
\texttt{alloc\_req}     & 1   & Tag Allocator & Pulse: allocate a new ROB slot \\
\texttt{AXID[5:0]}      & 6   & Tag Allocator & AXI ID of transaction being allocated \\
\texttt{seqnum[1:0]}    & 2   & Tag Allocator & Seqnum of transaction being allocated \\
\texttt{tag[7:0]}       & 8   & Tag Allocator & Full tag (used to write Lookup Table) \\
\texttt{RESP\_VALID}    & 1   & MC Core       & Completion valid strobe \\
\texttt{resp\_type[1:0]}& 2   & MC Core       & 00=RD\_DATA, 01=WR\_ACK, 10=ERR; ROB acts only on RD\_DATA \\
\texttt{tag[7:0]}       & 8   & MC Core       & Completion tag; \texttt{[7:2]}=AXID, \texttt{[1:0]}=seqnum \\
\texttt{data[511:0]}    & 512 & MC Core       & Read data payload \\
\texttt{status[3:0]}    & 4   & MC Core       & Completion status/error code \\
\bottomrule
\caption{ROB input signals}
\end{longtable}

\subsection*{Output Signals}
\begin{longtable}{llll}
\toprule
\textbf{Signal} & \textbf{Width} & \textbf{Destination} & \textbf{Description} \\
\midrule
\texttt{data[511:0]}    & 512 & Merge Logic  & In-order read data beat \\
\texttt{status[3:0]}    & 4   & Merge Logic  & Status for emitted beat \\
\texttt{AXID[5:0]}      & 6   & Merge Logic  & AXID of emitted beat \\
\texttt{rob\_last}      & 1   & Merge Logic  & Asserted on final beat of a transaction; sourced from RTT, passed through \\
\texttt{PKT\_READY}     & 1   & Packet FIFO  & Deasserted when watermark hit; halts new read requests \\
\texttt{RESP\_READY}    & 1   & MC Core      & Tied high permanently \\
\bottomrule
\caption{ROB output signals}
\end{longtable}

\section{Blocks}

\subsection{Free List}
\begin{itemize}[noitemsep]
  \item 64-bit bitmask. \texttt{1 = free}, \texttt{0 = occupied}.
  \item Allocate: priority encoder selects lowest free bit, clears it, outputs \texttt{free\_slot\_idx[5:0]}.
  \item Free: Emit Logic sets the bit at \texttt{slot\_idx} via \texttt{free\_req}.
  \item Outputs \texttt{free\_slot\_idx} to Lookup Table (write), ROB BRAM (init), and Head Pointer Array (first allocation).
\end{itemize}

\subsection{Occupancy Counter}
\begin{itemize}[noitemsep]
  \item Up/down counter, range 0--64.
  \item Incremented by \texttt{alloc\_req} from Free List.
  \item Decremented by \texttt{free\_req} from Emit Logic.
  \item Asserts \texttt{watermark\_hit} when count $\geq$ \texttt{MC\_RD\_BUF} $-$ \texttt{ROB\_WATERMARK} $= 38$.
\end{itemize}

\subsection{Lookup Table}
\begin{itemize}[noitemsep]
  \item 256-entry register file, addressed by \texttt{\{AXID[5:0], seqnum[1:0]\}} (8-bit address).
  \item Each entry stores \texttt{slot\_idx[5:0]}.
  \item Write port: Tag Allocator writes \texttt{slot\_idx} at \texttt{\{AXID, seqnum\}} on allocation.
  \item Read port: Completion Input reads \texttt{slot\_idx} using \texttt{\{tag[7:2], tag[1:0]\}}.
  \item Clear: Emit Logic clears entry at \texttt{\{AXID, seqnum\}} after emit.
\end{itemize}

\subsection{ROB BRAM}
\begin{itemize}[noitemsep]
  \item Dual-port BRAM, 64 $\times$ 517 bits per slot: \texttt{valid[0], data[511:0], status[3:0]}.
  \item Write port: Completion Input writes \texttt{data}, \texttt{status}, \texttt{valid=1} at \texttt{slot\_idx}.
  \item Write port (init): Free List writes \texttt{valid=0} at \texttt{free\_slot\_idx} on allocation.
  \item Read port: Head Pointer Array supplies \texttt{head[AXID]} as read address; all 64 head slots checked in parallel each cycle.
  \item 1-cycle read latency — Emit Logic must account for this in its valid check timing.
\end{itemize}

\subsection{Completion Input}
\begin{itemize}[noitemsep]
  \item Receives \texttt{RESP\_VALID, resp\_type[1:0], tag[7:0], data[511:0], status[3:0]}
        from MC Core.
  \item \textbf{Filters on \texttt{resp\_type}:} acts only when
        \texttt{resp\_type == 2'b00} (RD\_DATA). Completions with
        \texttt{resp\_type == 2'b01} (WR\_ACK) are destined for the Response Tracker
        and are ignored by the ROB. ERR completions (\texttt{resp\_type == 2'b10}) for
        read transactions are written into the ROB with the error status preserved so
        DECERR can be forwarded to the R Response Generator.
  \item Extracts \texttt{AXID = tag[7:2]}, \texttt{seqnum = tag[1:0]}.
  \item Reads \texttt{slot\_idx} from Lookup Table using \texttt{\{AXID, seqnum\}}.
  \item Writes \texttt{data}, \texttt{status}, \texttt{valid=1} to ROB BRAM at \texttt{slot\_idx}.
\end{itemize}

\subsection{Head Pointer Array}
\begin{itemize}[noitemsep]
  \item 64 registers of 6 bits each, one per AXID. Stores slot index of next expected beat.
  \item Written on first allocation only: \texttt{head[AXID] = free\_slot\_idx} when the
        AXID has no in-flight slots (occupancy transitions 0\textrightarrow{}1 for that AXID).
        On subsequent allocations for the same AXID, \texttt{head[AXID]} is \emph{not} updated
        — it is owned exclusively by Emit Logic from that point forward.
  \item Advanced by Emit Logic after each emit: incremented to the next expected slot index.
  \item Reset to the next \texttt{free\_slot\_idx} only when all in-flight slots for an AXID
        have been emitted and a new allocation arrives (occupancy returns to 0 then rises again).
  \item All 64 head pointers drive the ROB BRAM read port simultaneously each cycle.
\end{itemize}

\subsection{Emit Logic}
\begin{itemize}[noitemsep]
  \item Reads \texttt{valid} bits from all 64 head slots out of ROB BRAM each cycle.
  \item Priority encoder selects lowest AXID whose head slot has \texttt{valid=1}.
  \item On selection: reads \texttt{data} and \texttt{status} from BRAM; emits to Merge Logic
        along with \texttt{AXID} and \texttt{rob\_last}. \texttt{rob\_last} is \emph{not}
        generated inside the ROB — it is sourced from the Read Transaction Table (RTT), which
        asserts it when \texttt{completed\_packets == expected\_packets} for that transaction.
        The ROB passes it through to Merge Logic coincident with the final data beat.
  \item After emit:
    \begin{itemize}[noitemsep]
      \item Clears \texttt{valid} in BRAM slot.
      \item Advances \texttt{head[AXID]} in Head Pointer Array.
      \item Sends \texttt{free\_req + slot\_idx} to Free List (sets bitmask bit).
      \item Decrements Occupancy Counter via \texttt{free\_req}.
      \item Clears Lookup Table entry at \texttt{\{AXID, seqnum\}}.
    \end{itemize}
  \item Maximum one emission per cycle (single R channel path).
\end{itemize}

\subsection{PKT\_READY Control}
\begin{itemize}[noitemsep]
  \item Purely combinational.
  \item Asserts \texttt{PKT\_READY=0} to Packet FIFO when \texttt{watermark\_hit} is high.
  \item Halts new read requests to MC Core; prevents ROB overflow since \texttt{RESP\_READY} cannot be deasserted.
\end{itemize}

\section{Allocation Path}

\begin{enumerate}[noitemsep]
  \item Tag Allocator sends \texttt{alloc\_req} (pulse) with \texttt{AXID}, \texttt{seqnum}, \texttt{tag}.
  \item Free List pops \texttt{free\_slot\_idx} (priority encoder on bitmask), clears the bit.
  \item Free List writes \texttt{slot\_idx} to Lookup Table at \texttt{\{AXID, seqnum\}}.
  \item Free List initializes ROB BRAM slot: \texttt{valid=0} at \texttt{free\_slot\_idx}.
  \item Free List stores \texttt{free\_slot\_idx} into \texttt{head[AXID]} (first allocation for that AXID only).
  \item Occupancy Counter increments.
\end{enumerate}

\section{Completion Path}

\begin{enumerate}[noitemsep]
  \item MC Core asserts \texttt{RESP\_VALID} with \texttt{tag[7:0]}, \texttt{data[511:0]}, \texttt{status[3:0]}.
  \item Completion Input extracts \texttt{AXID = tag[7:2]}, \texttt{seqnum = tag[1:0]}.
  \item Completion Input reads \texttt{slot\_idx} from Lookup Table.
  \item Completion Input writes \texttt{data}, \texttt{status}, \texttt{valid=1} to ROB BRAM at \texttt{slot\_idx}.
  \item \texttt{RESP\_READY} is tied high — MC Core never stalled on this path.
\end{enumerate}

\section{Emit Path}

\begin{enumerate}[noitemsep]
  \item Each cycle, all 64 \texttt{head[AXID]} values drive the ROB BRAM read port.
  \item BRAM returns \texttt{valid} bits for all 64 head slots (1-cycle latency).
  \item Priority encoder in Emit Logic selects lowest AXID with \texttt{valid=1}.
  \item Emit Logic reads \texttt{data} and \texttt{status} from selected slot; forwards to Merge Logic with \texttt{AXID} and \texttt{last\_beat}.
  \item Slot is freed: \texttt{valid} cleared, \texttt{head[AXID]} advanced, Free List bit set, Occupancy Counter decremented, Lookup Table entry cleared.
\end{enumerate}

\section{Backpressure \& Overflow Prevention}

\begin{center}
\begin{tabularx}{\textwidth}{>{\ttfamily}l X}
\toprule
\textbf{Signal} & \textbf{Description} \\
\midrule
watermark\_hit  & Asserted when occupancy $\geq$ 38. Drives PKT\_READY control. \\
PKT\_READY=0    & Stops Packet FIFO from issuing new read requests to MC Core. \\
RESP\_READY=1   & Tied high permanently. ROB can never stall MC Core completions. \\
\bottomrule
\end{tabularx}
\end{center}

\medskip
\noindent Watermark headroom = \texttt{ROB\_WATERMARK = 26}. Ensures all requests already accepted by MC Core can return completions without overflowing the ROB.

\section{Key Parameters}

\begin{center}
\begin{tabular}{ll}
\toprule
\textbf{Parameter} & \textbf{Value} \\
\midrule
ROB depth         & 64 entries (\texttt{MC\_RD\_BUF}) \\
Slot width        & 517 bits (1 valid + 512 data + 4 status) \\
Free list width   & 64 bits \\
Lookup table      & 256 entries $\times$ 6-bit slot index \\
Head pointers     & 64 $\times$ 6-bit registers \\
Watermark         & 38 (64 $-$ 26) \\
Max emissions     & 1 per cycle \\
\bottomrule
\end{tabular}
\end{center}

\section{Timing Notes}

\begin{itemize}[noitemsep]
  \item ROB BRAM has 1-cycle read latency. Emit Logic valid check must be pipelined accordingly — register the read address, use valid bit one cycle after presenting \texttt{head[AXID]}.
  \item Priority encoder across 64 valid bits is the critical path in Emit Logic. May need pipelining at high frequencies.
  \item Allocation and completion can occur in the same cycle — write port (completion) and init write (allocation) are independent BRAM ports.
  \item \texttt{watermark\_hit} is combinational from the Occupancy Counter — no additional
        latency before \texttt{PKT\_READY} deasserts.
  \item Simultaneous \texttt{alloc\_req} and \texttt{free\_req} on the Free List bitmask in
        the same cycle is safe — they always target different bits (alloc clears the lowest
        free bit; free sets the bit of the just-emitted slot, which is by definition occupied).
        No read-modify-write conflict arises.
\end{itemize}

\end{document}
